The Heavy Photon Search Test Detector

### 2014-04-03 hps-testrun-nim.tex
Comments by Matt Graham

\section{Performance}

@@

After a trigger is received, the amplitude  of every  APV25 analoge output  is sampled and digitized in 
six consecutive time bins, separated by roughly 25~ns.  
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape}. As the figure demonstrates,  
the SVT was well timed-in to the trigger with the rise of the pulse at the 3rd sampling 
point. Note that in the Test run the APV25 ASICs were operating with a 50~ns shaping time.  
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{Accumulation of six pedestal-subtracted samples from individual SVT channels associated 
	with hits on tracks.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}
These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips with the position of a cluster on the sensor determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape. 
%, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
%\begin{figure}[]
%\begin{center}
%{\small
%	%\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
%	\includegraphics[width=7cm]{figures/mip_top_layer_2.png}
%    	\caption{ The cluster charge distribution for hits associated with a track. }
%	\label{fig:cluster_pulse}
%}
%\end{center}
%\end{figure}

In order to find the time, $t_0$, and amplitude of each hit, the six samples from each channel are fitted 
to an ideal $CR-RC$ function. After clustering hits on a sensor, the hit time for each cluster is computed 
as the amplitude-weighted average of the individually fitted $t_0$ on each channel. The 
$t_0$-resolution is studied by comparing the cluster hit time with the average of all cluster hit times on 
the track, see Fig.~\ref{fig:tracktime}. 
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/test_run_1351_hit_time_corrected_top_layer2_mod.pdf}
	\caption{
	The residual of individual cluster times with the average of all clusters on the track. }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected for S/N~25 which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 ASIC pulse shaping time will also improve time resolution. 
These results show that we can operate with the six sample readout mode of the APV25 chip and 
achieve time resolution adequate for pileup rejection during electron running in HPS.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency was estimated by measuring 
the number of good tracks with a hit close to the extrapolated intersection of a given sensor that was 
excluded from the track fit itself. Tracks which intersect regions with known bad channels or very close to 
in the edge region are excluded. The hit efficiency, see Fig.~\ref{fig:hit_efficiency}, was measured to be 
above 98\% and fairly uniform across the SVT.
\begin{figure}[]
\begin{center}
{\small
    	\includegraphics[width=7cm]{figures/single_hit_efficiency_Omar_11192013.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}
The spatial resolution of similar microstrip sensors is well established by test beam data, against which 
the charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ratio above 20, as demonstrated 
here, is relatively constant at approximately 6~$\mu$m for tracks that enters approximately normal to the 
sensors as in HPS.

@@

The expected momentum resolution from simulation is between 4-5\% for tracks in the momentum 
range of the Test run. By looking at the agreement 
between data and simulation in the shape of the kinematic distributions for single- and two track 
events and adjusting the overall momentum scale and smearing simulated events with a 
Gaussian resolution we estimate an agreement with the nominal scale and resolution to within 10\%.
 \begin{center}
{\small
\begin{figure*}[t]
   \includegraphics[width=5cm]{figures/h_trk_top_px_h_trk_top_px_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_invM_h_invM_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_sumE_h_sumE_trigsel4hit_pair1351_twotrkfilt-v6-paper}
\caption{Kinematic distributions for e$^+$e$^-$ pairs selected by opposite charged tracks in the top and bottom half of the tracker: track momentum in the top half of the SVT (left), invariant mass (middle) and the sum of the track momentum for the pair (right).}
\label{fig:pair_kin}
\end{figure*}
}
\end{center}
In the Test run, as well as in electron running with HPS, the dominant source of uncertainty in the 
tracking and vertexing is multiple Coulomb scattering. For the vertexing performance the foremost 
difference compared to electron beam running is that the target was located approximately 67~cm 
upstream from our nominal target position; giving almost collinear tracks in the detector. The increased 
lever arm over which tracks are extrapolated widens the resolution with up to a factor of eight (depending 
on momentum) compared to what is achieved at the nominal electron target position for HPS. 
Figure~\ref{fig:impact_param} shows the horizontal and vertical positions of the extrapolated track at the 
converter position. 
{\small
	 \begin{figure}[]
 \begin{center}
	% Plot info:
	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
	\includegraphics[width=7cm]{figures/h_trk_top_fr_conv_z_h_trk_top_conv_z_dataMC_twotrksel.png}
	\includegraphics[width=7cm]{figures/h_trk_top_fr_conv_y_h_trk_top_conv_y_dataMC_twotrksel.png}
	\caption{ Vertical (top) and horizontal (bottom) extrapolated track position at the converter position 
	taking into account the measured fringe field. }
	\label{fig:impact_param}
\end{center}
\end{figure}
}
While residual alignment show small shifts good agreement on the widths between data and simulated 
events, indicating a good understanding of our material budget and distribution in the SVT.  Having 
the dominant contribution to the vertex resolution approximately right demonstrates that the resolution 
in HPS, with a target at 10~cm, will be as calculated.
%Furthermore, tails of the vertex distributions are impossible to study with the finite data sample obtained 
%in the Test run. Nevertheless, useful information can still be 
%obtained by studying the vertex distributions. Figure~\ref{fig:vtx_pos} shows the distance of closest 
%approach of the momentum vectors extrapolated in the upstream direction from our analyzing magnet, 
%taking into account the measured fringe field of the PS dipole magnet. 
 %\begin{center}
%{\small
%	 \begin{figure*}[t]
%	% Plot info:
%	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
%	\includegraphics[width=5cm]{figures/h_vtx_fr_x_h_vtx_x_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\includegraphics[width=5cm]{figures/h_vtx_fr_y_h_vtx_y_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\includegraphics[width=5cm]{figures/h_vtx_fr_z_h_vtx_z_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\caption{ Vertex position represented by the distance of closest approach of the extrapolated 
%	momentum vectors upstream of the PS dipole magnet taking into account the measured fringe field. 
%	The overall shift from zero in the x-direction is due to a 30~mrad rotation of the SVT with respect to 
%	the beam line.}
%	\label{fig:vtx_pos}
%\end{figure*}
%}
%\end{center}

@@

The ratio of the ECal cluster energy $E$ to the momentum $p$ of a matched track in the SVT was used 
to determine the conversion factors from ADC counts to energy. To compare data and simulation, all 
inoperable or noisy channels in the SVT and ECal was disabled in both data and simulation so that any 
efficiency or bias that affect the data should be reflected in the simulation. 
%Furthermore, a conversion factor calibration is performed to equalize the gains of  each channel. 
%simulation to those observed in data by comparing .  The ``weighted $E/p$'' for each crystal is %computed, 
%representing the average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, conversion coefficients for each crystal is adjusted until the $E$/$p$ ratio in data and 
simulation are similar. The distribution of the $E$/$p$ ratio in data and simulation are compared in 
Fig.~\ref{fig:gains}. 
{\small
\begin{figure}[]
\begin{center}
	\includegraphics[width=7cm]{figures/h_ep_data_0_h_ep_MC_0_dataMC_1351-v6-v6gains_2-trig-top-cl600reg0.pdf}
	\caption{The ECal energy over track momentum ratio ($E$/$p$) comparing data and simulation 
	for inclusive events in the top half of the ECal.} 
	\label{fig:gains}
\end{center}
\end{figure}
}
The peak position of the distribution indicates the sampling fraction of the ECal, the fraction of the 
incident particle energy measured in the cluster. The width and tails of the distribution 
in data indicates imperfect calibration and noise of the ECal modules. This level of calibration and the 
agreement with simulation was found to be sufficient to study normalized event rates on the Test run.

\subsection{Trigger Performance}
As described above the energy from each 
crystal is measured differently in the trigger and what is readout from the ECal. 
The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
To eliminate trigger bias, we use a tag and probe method: to study the trigger performance in one half 
of the ECal, we select 
events which triggered the other half and where there was exactly one probe cluster in the ECal half 
under study. We then measure trigger efficiency as the fraction of tagged events that fired the trigger in 
the probe half as a function of the probe cluster energy, shown in Fig.~\ref{fig:turnon}. 
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/h_cl_E_probedata_eff_h_cl_E_probeMC_eff_dataMC_1351-v7-trig-tagBot.pdf}
	\caption{Trigger efficiency in both halves of the ECal for data and simulation as a 
	function of cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below 1~GeV for two reasons;  
gain variations between different crystals lead to the threshold variations and the nonlinearity of 
the time-over-threshold integral means that the effective threshold is higher for clusters that span 
multiple crystals. The effective trigger threshold is therefore dependent on position and energy of 
the particle as well as cluster multiplicity.

@@

\subsection{Trigger Rate Comparisons}
\label{sec:mcs}
Trigger rates observed in the Test run are dominated by multiple Coulomb scattered \ee pairs in the 
converter. In simulated events, the rate of triggers depend on the modeling of the pairs angular 
distribution and the subsequent multiple Coulomb scattering in the converter. Rates from different 
converter 
thicknesses are used to study the varying multiple Coulomb scattering contribution (pair production 
angle is constant). Restricting clusters to a well calibrated region of the ECal and subtracting the 
"no converter" background we see agreement with the rates predicted by the \egs{} simulation 
program, see Tab.~\ref{tab:mcs}.
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
GEANT4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted number of events for 1~s of beam at 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.}
}
\label{tab:mcs}
\end{table}
%\end{center}
This gives further confidence that the dominant source of background occupancy for HPS, multiple 
Coulomb scattered beam electrons, is well 
described~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467} . 
%One of the main goals of the Test run was to evaluate the description of the tails of the 
%multiple Coulomb scattering to gain further confidence in the expected detector occupancy in the full HPS 
%experiment. Those occupancies have been evaluated using \egs{}~\cite{egs5} simulation program 
%whose \moliere{} formulation of multiple scattering show good agreement with previous studies 
%for a wide range of materials and projectiles~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467}. 
%Despite all data taken with a photon beam in the Test run,

>>>>

\section{Reconstruction and Performance}

@@

After a trigger is received, the amplitude  of every  APV25 analogue output is sampled and digitized in 
six consecutive time bins, separated by roughly 25~ns.  
The typical, pedestal subtracted, pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape}. 
As the figure demonstrates,  
the SVT was well timed-in to the trigger with the rise of the pulse at the 3rd sampling 
point.  
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{Accumulation of six pedestal-subtracted samples from individual SVT channels associated 
	with hits on tracks.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}
In order to find the time, $t_0$, and amplitude of each hit, the six samples from each channel are fitted 
to an ideal $CR-RC$ function. Note that in the Test run the APV25 ASICs were operating with a 50~ns 
shaping time. 
These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips with the position of a cluster on the sensor determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape. 
%, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[]
\begin{center}
{\small
	%\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
	\includegraphics[width=7cm]{figures/mip_top_layer_2.png}
    	\caption{ The cluster charge distribution for hits associated with track follow the characteristic 		Landau shape. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}
After clustering hits on a sensor, the hit time for each cluster is computed 
as the amplitude-weighted average of the individually fitted $t_0$ on each channel. The 
$t_0$ resolution is studied by comparing the cluster hit time with the average of all cluster hit times on 
the track shown in Fig.~\ref{fig:tracktime}. 
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/test_run_1351_hit_time_corrected_top_layer2_mod.pdf}
	\caption{
	The residual of individual cluster times with the average of all clusters on the track. }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected for S/N~25 which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 ASIC pulse shaping time to 35~ns will also improve time resolution. 
These results show that we can operate with the six sample readout mode of the APV25 chip and 
achieve time resolution adequate for pileup rejection during electron running in HPS.

Good agreement between observed and simulated occupancies after taking into account dead or 
noisy channels. The hit reconstruction efficiency was estimated by measuring 
the number of good tracks with a hit close to the extrapolated intersection of a given sensor that was 
excluded from the track fit itself. Tracks which intersect regions with known bad channels or very close 
to in the edge region are excluded. The hit reconstruction efficiency, see Fig.~\ref{fig:hit_efficiency}, 
was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[]
\begin{center}
{\small
    	\includegraphics[width=7cm]{figures/single_hit_efficiency_Omar_11192013.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}

The spatial resolution of similar microstrip sensors is well established by test beam data, against which 
the charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ratio above 20, as demonstrated 
here, is relatively constant at approximately 6~$\mu$m for tracks that enters approximately normal to the 
sensors as in HPS.

@@

The expected momentum resolution from simulation is between 4-5\% for tracks in the momentum 
range of the Test run.

By comparing the shapes of the kinematic distributions for single- and two track events between data 
and simulation we estimate an agreement with the nominal scale and resolution to within 10\%.
 \begin{center}
{\small
\begin{figure*}[t]
   \includegraphics[width=5cm]{figures/h_trk_top_px_h_trk_top_px_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_invM_h_invM_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_sumE_h_sumE_trigsel4hit_pair1351_twotrkfilt-v6-paper}
\caption{Kinematic distributions for e$^+$e$^-$ pairs selected by opposite charged tracks in the top and bottom half of the tracker: track momentum in the top half of the SVT (left), invariant mass (middle) and the sum of the track momentum for the pair (right).}
\label{fig:pair_kin}
\end{figure*}
}
\end{center}
In the Test run, as well as in electron running with HPS, the dominant source of uncertainty in the 
tracking and vertexing is multiple Coulomb scattering. For the vertexing performance the foremost 
difference compared to electron beam running is that the target was located approximately 67~cm 
upstream from our nominal target position; giving almost collinear tracks in the detector. The increased 
lever arm over which tracks are extrapolated widens the resolution with up to a factor of eight (depending 
on momentum) compared to what is achieved at the nominal electron target position for HPS. 
Figure~\ref{fig:impact_param} shows the horizontal and vertical positions of the extrapolated track at the 
converter position. 
{\small
	 \begin{figure}[]
 \begin{center}
	% Plot info:
	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
	\includegraphics[width=7cm]{figures/h_trk_top_fr_conv_z_h_trk_top_conv_z_dataMC_twotrksel.png}
	\includegraphics[width=7cm]{figures/h_trk_top_fr_conv_y_h_trk_top_conv_y_dataMC_twotrksel.png}
	\caption{ Vertical (top) and horizontal (bottom) extrapolated track position at the converter position 
	taking into account the measured fringe field. }
	\label{fig:impact_param}
\end{center}
\end{figure}
}
While residual alignment show small shifts, the good agreement  between data and simulated 
events of the widths indicates a good understanding of the material budget and distribution in the SVT.  
Having  the dominant contribution to the vertex resolution approximately right demonstrates that the 
resolution in HPS, with a target at 10~cm, will be as calculated.
%Furthermore, tails of the vertex distributions are impossible to study with the finite data sample obtained 
%in the Test run. Nevertheless, useful information can still be 
%obtained by studying the vertex distributions. Figure~\ref{fig:vtx_pos} shows the distance of closest 
%approach of the momentum vectors extrapolated in the upstream direction from our analyzing magnet, 
%taking into account the measured fringe field of the PS dipole magnet. 
 %\begin{center}
%{\small
%	 \begin{figure*}[t]
%	% Plot info:
%	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
%	\includegraphics[width=5cm]{figures/h_vtx_fr_x_h_vtx_x_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\includegraphics[width=5cm]{figures/h_vtx_fr_y_h_vtx_y_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\includegraphics[width=5cm]{figures/h_vtx_fr_z_h_vtx_z_trigsel4hit_pair1351_twotrkfilt-v6-paper}
%	\caption{ Vertex position represented by the distance of closest approach of the extrapolated 
%	momentum vectors upstream of the PS dipole magnet taking into account the measured fringe field. 
%	The overall shift from zero in the x-direction is due to a 30~mrad rotation of the SVT with respect to 
%	the beam line.}
%	\label{fig:vtx_pos}
%\end{figure*}
%}
%\end{center}

@@

The ratio of the ECal cluster energy $E$ to the momentum $p$ of a matched track in the SVT was used 
to determine the conversion factors from ADC counts to energy. To compare data and simulation, all 
inoperable or noisy channels in the SVT and ECal were disabled in both data and simulation so that 
any efficiency or bias that affect the data should be reflected in the simulation. 
%Furthermore, a conversion factor calibration is performed to equalize the gains of  each channel. 
%simulation to those observed in data by comparing .  The ``weighted $E/p$'' for each crystal is %computed, 
%representing the average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, conversion coefficients for each crystal were adjusted until the $E$/$p$ ratio in data and 
simulation were similar. The distribution of the $E$/$p$ ratio in data and simulation are compared in 
Fig.~\ref{fig:gains}. 
{\small
\begin{figure}[]
\begin{center}
	\includegraphics[width=7cm]{figures/h_ep_data_0_h_ep_MC_0_dataMC_1351-v6-v6gains_2-trig-top-cl600reg0.pdf}
	\caption{The ECal energy over track momentum ratio ($E$/$p$) comparing data and simulation 
	for single cluster triggers in the top half of the ECal.} 
	\label{fig:gains}
\end{center}
\end{figure}
}
The peak position of the distribution indicates the sampling fraction of the ECal, the fraction of the 
incident particle energy measured in the cluster. The width and tails of the distribution 
in data indicates imperfect calibration and noise of the ECal modules. This level of calibration and the 
agreement with simulation was found to be sufficient to study normalized event rates on the Test run.

\subsection{Trigger Performance}
As described above in Sec.~\ref{sec:triggerdaq}, the energy from each 
crystal is measured differently in the trigger and what is readout from the ECal. 
The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
To eliminate trigger bias, we use a tag and probe method: to study the trigger performance in one half 
of the ECal, we select 
events which triggered the other half and where there was exactly one probe cluster in the ECal half 
under study. We then measure trigger efficiency as the fraction of tagged events that fired the trigger in 
the probe half as a function of the probe cluster energy, shown in Fig.~\ref{fig:turnon}. 
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/h_cl_E_probedata_eff_h_cl_E_probeMC_eff_dataMC_1351-v7-trig-tagBot.pdf}
	\caption{Trigger efficiency in both halves of the ECal for data and simulation as a 
	function of cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below 1~GeV for two reasons;  
gain variations between different crystals lead to the threshold variations and the nonlinearity of 
the time-over-threshold integral means that the effective threshold is higher for clusters that span 
multiple crystals. The effective trigger threshold is therefore dependent on position and energy of 
the particle as well as cluster multiplicity.

@@

\subsection{Trigger Rate Comparisons}
\label{sec:mcs}
Trigger rates observed in the Test run are dominated by multiple Coulomb scattered \ee{} pairs in the 
converter. In simulated events, the rate of triggers depend on the modeling of the pairs angular 
distribution and the subsequent multiple Coulomb scattering in the converter. Rates from different 
converter 
thicknesses are used to study the varying multiple Coulomb scattering contribution (pair production 
angle is constant). Restricting clusters to a well calibrated region of the ECal and subtracting the 
"no converter" background we see agreement with the rates predicted by the \egs{} simulation 
program, see Tab.~\ref{tab:mcs}.
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
GEANT4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted event rate (in Hz) normalized to 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.}
}
\label{tab:mcs}
\end{table}
%\end{center}
This gives further confidence that the dominant source of background occupancy for HPS, multiple 
Coulomb scattered beam electrons, is well 
described~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467} . 
%One of the main goals of the Test run was to evaluate the description of the tails of the 
%multiple Coulomb scattering to gain further confidence in the expected detector occupancy in the full HPS 
%experiment. Those occupancies have been evaluated using \egs{}~\cite{egs5} simulation program 
%whose \moliere{} formulation of multiple scattering show good agreement with previous studies 
%for a wide range of materials and projectiles~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467}. 
%Despite all data taken with a photon beam in the Test run,
